Performing and evaluating graph-based clustering with rOpenScPCA

### 2024-11-13 analyses/hello-clusters/01_perform-evaluate-clustering.Rmd
response to reviews: add parentheses for functions, just use backticks for pca names, and add missing wording

---
title: "Performing and evaluating graph-based clustering with rOpenScPCA"
date: "`r Sys.Date()`"
author: Data Lab
params:
  seed: 2024
output: 
  html_notebook: 
    toc: yes
    toc_float: yes
    df_print: paged
---

@@

This section will show how to perform clustering with the function `rOpenScPCA::calculate_clusters`.

@@

* The `algorithm` can be one of:
  * `"louvain"`, `"walktrap"`, or `"leiden"`
* The `weighting` can be one of:
  * `"jaccard"`, `"rank"`, or `"number"`
* The nearest neighbors parameter can be customized with the `nn` argument
* The resolution parameter can be customized with the `resolution` argument
  * This parameter is only used by Louvain and Leiden algorithms
* If the Leiden algorithm is used, its default objective function parameter will be `"CPM"`, but you can also set  `objective_function = "modularity"` instead.
* You can provide additional parameters as a list to the `cluster_args` argument.
  * Please refer to the [`igraph` documentation](https://igraph.org/r/html/latest) to learn more about what additional parameters can be provided to each clustering algorithm.
  * Note that `cluster_args` only accepts single-length arguments (no vectors or lists).

@@

When performing clustering on an SCE object, `rOpenScPCA::calculate_clusters` assumes that the principal components are stored in a reduced dimension slot named `"PCA"`.
If your 
The function assumes the PC name is `"PCA"`. 
If your SCE has a different name for the principal components, you can provide the name with the argument `pc_name`.

@@

When performing clustering on a Seurat object, `rOpenScPCA::calculate_clusters` assumes that the principal components are stored in a reduced dimension slot named `"pca"`.
If your Seurat has a different name for the principal components, you can provide the name with the argument `pc_name`.
In this example, because we created the Seurat object from the SCE object, its principal components are named `"PCA"`, so we will provide that information.

@@

```{r cluster seurat}
cluster_results <- calculate_clusters(
  seurat_obj,
  # Since we converted this from an SCE object, its PC name is `"PCA"`
  pc_name = "PCA",
  seed = params$seed
)

@@

* An SCE or Seurat object that contains principal components
  * If the PCA matrix name is not the default for the object type (`"PCA"` for SCE and `"pca"` for Seurat), you can use `pc_name` to specify its name.
* A data frame with, at least, columns representing unique cell ids and cluster assignments.
  * By default, these columns should be named `cell_id` and `cluster`, respectively, matching the output of `rOpenScPCA::calculate_clusters`
  * You can override these defaults using the arguments `cell_id_col` and `cluster_col`

@@

We'll use the function `rOpenScPCA::calculate_silhouette` to calculate the silhouette width, and then we'll visualize the results.

@@

We'll use the function `rOpenScPCA::calculate_purity` to calculate the neighborhood purity, and then visualize the results.

@@

We'll use the function `rOpenScPCA::calculate_stability` to calculate the stability of clusters using a bootstrapping approach, and then visualize the results.
By default, this function performs 20 bootstrap replicates, but this can be customized using the argument `replicates`.

@@

When calculating bootstrap clusters, `rOpenScPCA::calculate_stability` uses `rOpenScPCA::calculate_clusters` with default parameters. 
If your original clusters were not calculated with these defaults, you should pass those customized values into this function as well to ensure a fair comparison between your original clusters and the bootstrap clusters.

@@

Results can either be directly exported as a TSV file (e.g., with `readr::write_tsv`), or you can add the results into your scRNA-seq object.
The subsequent examples will show you, as an example, how to save the `cluster_results` data frame to a given object.

>>>>

---
title: "Performing and evaluating graph-based clustering with rOpenScPCA"
date: `r Sys.Date()`
author: Data Lab
params:
  seed: 2024
output: 
  html_notebook: 
    toc: yes
    toc_float: yes
    df_print: paged
---

@@

This section will show how to perform clustering with the function `rOpenScPCA::calculate_clusters()`.

@@

* The `algorithm` can be one of:
  * `louvain`, `walktrap`, or `leiden`
* The `weighting` can be one of:
  * `jaccard`, `rank`, or `number`
* The nearest neighbors parameter can be customized with the `nn` argument
* The resolution parameter can be customized with the `resolution` argument
  * This parameter is only used by Louvain and Leiden algorithms
* If the Leiden algorithm is used, its default objective function parameter will be `CPM`, but you can also set  `objective_function = "modularity` instead.
* You can provide additional parameters as a list to the `cluster_args` argument.
  * Please refer to the [`igraph` documentation](https://igraph.org/r/html/latest) to learn more about what additional parameters can be provided to each clustering algorithm.
  * Note that `cluster_args` only accepts single-length arguments (no vectors or lists).

@@

When performing clustering on an SCE object, `rOpenScPCA::calculate_clusters()` assumes that the principal components are stored in a reduced dimension slot named `PCA`.
The function assumes the PC name is `PCA`. 
If your SCE object has a different name for the principal components, you can provide the name with the argument `pc_name`.

@@

When performing clustering on a Seurat object, `rOpenScPCA::calculate_clusters()` assumes that the principal components are stored in a reduced dimension slot named `pca`.
If your Seurat object has a different name for the principal components, you can provide the name with the argument `pc_name`.
In this example, because we created the Seurat object from the SCE object, its principal components are named `PCA`, so we will provide that information.

@@

```{r cluster seurat}
cluster_results <- calculate_clusters(
  seurat_obj,
  # Since we converted this from an SCE object, its PC name is `PCA`
  pc_name = "PCA",
  seed = params$seed
)

@@

* An SCE or Seurat object that contains principal components
  * If the PCA matrix name is not the default for the object type (`PCA` for SCE and `pca` for Seurat), you can use `pc_name` to specify its name.
* A data frame with, at least, columns representing unique cell ids and cluster assignments.
  * By default, these columns should be named `cell_id` and `cluster`, respectively, matching the output of `rOpenScPCA::calculate_clusters()`
  * You can override these defaults using the arguments `cell_id_col` and `cluster_col`

@@

We'll use the function `rOpenScPCA::calculate_silhouette()` to calculate the silhouette width, and then we'll visualize the results.

@@

We'll use the function `rOpenScPCA::calculate_purity()` to calculate the neighborhood purity, and then visualize the results.

@@

We'll use the function `rOpenScPCA::calculate_stability()` to calculate the stability of clusters using a bootstrapping approach, and then visualize the results.
By default, this function performs 20 bootstrap replicates, but this can be customized using the argument `replicates`.

@@

When calculating bootstrap clusters, `rOpenScPCA::calculate_stability()` uses `rOpenScPCA::calculate_clusters()` with default parameters. 
If your original clusters were not calculated with these defaults, you should pass those customized values into this function as well to ensure a fair comparison between your original clusters and the bootstrap clusters.

@@

Results can either be directly exported as a TSV file (e.g., with `readr::write_tsv()`), or you can add the results into your scRNA-seq object.
The subsequent examples will show you, as an example, how to save the `cluster_results` data frame to a given object.
